---
title: 'Part 1: Dimensionality Reduction'
author: "VICTOR"
date: "6/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#    Dimensionality Reduction
## a) Data Analytic Question

The aim of  this project is to reduce the dataset to a low dimensional dataset via the t-SNE algorithm or PCA.

## b) Success Metrics

* Successful Loading the data.
* Successful Handling missing data.
* Successful Outliers detection.
* Successful Outlier Visualization.
* Successful Handling  outliers.
* Successful Univariate analysis.
* Successful Bivariate analysis.



## c) Context

undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest number of sales.

## d) Data Understanding
Variables

* The dataset consists of 8 numerical and 8 categorical attributes.
* Invoice.ID
* Branch
* Customer.type
* Gender
* Product.line
* Unit.price
* Quantity
* Tax
* Date
* Time
* Payment
* cogs
* gross.margin.percentage
* gross.income
* Rating
* Total

## e) Experimental Design

* Formulation of the research question.
* Data Sourcing
* Check the Data
* Perform Data Cleaning
* Perform Exploratory Data Analysis  (Univariate, Bivariate & Multivariate)
* Implement the Solution
* Challenging the Solution
* Follow up Questions

```{r cleaning,include=FALSE}

###############cleaning r ENVIROMENT#####################################################3

ls()  # TO see the objects you have created.
rm(list=ls()) #First delete all the objects using rm(list=ls())
gc()    #Then clear any occupied memory by running garbage collector using gc().
#############################################################################
library(AggregateR)
library(broom)
library(data.table)
library(date,anytime)
library(deSolve)
library(distcrete)
library(dplyr)
library(DT)
library(earlyR)
library(EpiEstim)
library(epitrix)
library(ff)
library(ggforce)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(gt)
library(Hmisc)
library(hrbrthemes)
library(incidence)
library(knitr)
library(lubridate)
library(magrittr)
library(projections)
library(readr)
library(rvest)
library(scales)
library(stringr)
library(tibble)
library(tidyverse)
library(writexl)
library(xtable)
library(class)
library(caTools)
library(psych)
library(ISOcodes)
library(caret)
library(countrycode)
library(mlbench)
library(e1071)
library(ggcorrplot)
options( java.parameters = "-Xmx4g")
options(digits = 15)

```


## Data Importation
```{r  data, include=TRUE}
dataset1<- read.csv("http://bit.ly/CarreFourDataset",header =T)


```


## converting data.frame data into data.table

```{r  converting, include=TRUE}

dataset1<-as.data.table(dataset1)
class(dataset1) #checking class

```

## Data Columns
```{r  preview, include=TRUE}

kable(colnames(dataset1))

```

## Check for missing values

```{r  missing, include=TRUE}
library(Amelia)
missmap(dataset1,main="Missing Values in Data Set")
#colSums(is.na(dataset1))

```

## any NAs in data set?

```{r  NAs, include=TRUE}
colSums(is.na(dataset1))

```

Now lets find the duplicated rows in the dataset df and assign to a variable duplicated_rows below.

```{r  duplicate, include=TRUE}
duplicated_rows <- dataset1[duplicated(dataset1),]
#Lets print out the variable duplicated_rows and see these duplicated rows
#kable(duplicated_rows)
```


Removing these duplicated rows in the data set or 
showing these unique items and assigning to a variable unique_items below
 
```{r  unique_items, include=TRUE}

unique_items <- dataset1[!duplicated(dataset1), ]

```
# Drop unnecessary column
 
```{r drop, include=TRUE}

dataset1 <- subset( dataset1, select = -Invoice.ID )
```

## Outlier Treatment

```{r  Outlier, include=TRUE}

mod <- lm( gross.margin.percentage~gross.income, data=dataset1)
cooksd <- cooks.distance(mod)

#Influence measures
#In general use, those observations that have a cook’s distance greater than 4 times 
#the mean may be classified as Outlier


plot(cooksd, pch="*", cex=2, main="Outliers by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels

```

## Tibbles


A tibble is a special kind of data.frame used by dplyr and other packages of the tidyverse. Tidyverse is a set of packages for data science that work in harmony because they share common data representations and API design. When a data.frame is turned into a tibble its class will change.

```{r  Tibbles, include=TRUE}

class(dataset1)

dataset1<- tbl_df(dataset1)

class(dataset1)


```


## Data Overview

```{r  Glimpse, include=TRUE,echo = FALSE}

glimpse(dataset1)
```

## Number of columns
```{r  columns, include=TRUE,echo = FALSE}
length(dataset1)
```

## Dimesion
```{r  Dimesion, include=TRUE,echo = FALSE}
dim(dataset1)

```

## Columnames
```{r  Columnames, include=TRUE,echo = FALSE}
colnames(dataset1)

```
# Encoding Categorical Variables
```{r  Encoding_m, include=TRUE,echo = FALSE}
library(encode)

dataset1$Branch<-as.factor(dataset1$Branch)
dataset1$Branch<-unclass(dataset1$Branch)         # Convert categorical variables
dataset1$Branch<- as.numeric(dataset1$Branch)


dataset1$Customer.type<-as.factor(dataset1$Customer.type)
dataset1$Customer.type<-unclass(dataset1$Customer.type) 
dataset1$Customer.type<- as.numeric(dataset1$Customer.type)


dataset1$Gender<-as.factor(dataset1$Gender)
dataset1$Gender<-unclass(dataset1$Gender)
dataset1$Gender<- as.numeric(dataset1$Gender)

dataset1$Product.line<-as.factor(dataset1$Product.line)
dataset1$Product.line<-unclass(dataset1$Product.line) 
dataset1$Product.line<- as.numeric(dataset1$Product.line)


dataset1$Payment<-as.factor(dataset1$Payment)
dataset1$Payment<-unclass(dataset1$Payment)
dataset1$Payment<- as.numeric(dataset1$Payment)

dataset1$Date<-as.Date(dataset1$Date,format='%m/%d/%Y')
dataset1$Date<-mday(dataset1$Date)
dataset1$Date<- as.numeric(dataset1$Date)

dataset1$Time<-as.POSIXct(dataset1$Time,format="%H:%M")
dataset1$Time<-hour(dataset1$Time)
dataset1$Time<- as.numeric(dataset1$Time)

```

## Change data types 

```{r  scale, include=TRUE,echo = FALSE}
#library(caret)
#dataset1$Quantity<- as.numeric(dataset1$Quantity)
#dataset1$Branch<- as.numeric(dataset1$Branch)
#dataset1$Customer.type<- as.numeric(dataset1$Customer.type)
#dataset1$Gender<- as.numeric(dataset1$Gender)
#dataset1$Product.line<- as.numeric(dataset1$Product.line)
#dataset1$Date<- as.numeric(dataset1$Date)
#dataset1$Time<- as.numeric(dataset1$Time)
#dataset1$Payment<- as.numeric(dataset1$Payment)
```

## Column data types

```{r  data_types, include=TRUE,echo = FALSE}
sapply(dataset1,class)
```

# UNIVARIATE ANALYSIS

## Unit.price
```{r rev price, include=TRUE,echo = FALSE}
   hist(dataset1$Unit.price,border="blue",col="green")
"mean"
mean(dataset1$Unit.price,na.rm=TRUE)

"median"

median(dataset1$Unit.price,na.rm=TRUE)

"mode"
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
getmode(dataset1$Unit.price)

   
```
## Tax
```{r Tax, include=TRUE,echo = FALSE}
   hist(dataset1$Tax,border="blue",col="cyan") 
"mean"
mean(dataset1$Tax,na.rm=TRUE)

"median"

median(dataset1$Tax,na.rm=TRUE)

"mode"
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
getmode(dataset1$Tax)

```
## Total
```{r  g.Total, include=TRUE,echo = FALSE}
   hist(dataset1$Total,border="blue",col="purple") 
"mean"
mean(dataset1$Total,na.rm=TRUE)

"median"

median(dataset1$Total ,na.rm=TRUE)

"mode"
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
getmode(dataset1$Total)

```


##  gross.income
```{r g.income, include=TRUE,echo = FALSE}
   hist(dataset1$gross.income,border="blue",col="deeppink") 
"mean"
mean(dataset1$gross.income,na.rm=TRUE)

"median"

median(dataset1$gross.income,na.rm=TRUE)

"mode"
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
getmode(dataset1$gross.income)

```

##  Time
```{r Time, include=TRUE,echo = FALSE}
   hist(dataset1$Time,border="blue",col="orange") 
"mean"
mean(dataset1$Time,na.rm=TRUE)

"median"

median(dataset1$Time,na.rm=TRUE)

"mode"
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
getmode(dataset1$Time)

```

##  Date
```{r Date, include=TRUE,echo = FALSE}
   hist(dataset1$Date,border="blue",col="navyblue") 
"median"

median(dataset1$Date,na.rm=TRUE)

"mode"
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
getmode(dataset1$Date)

```

# Correlation Matrix 

```{r  Correlation , include=TRUE,echo = FALSE}
#library(ggcorrplot)
#library(ggplot2)

#corrdata<-dataset1[,c("Unit.price","Quantity","Tax","gross.margin.percentage","gross.income","Rating","Total")]  
#corrdata<-na.omit(corrdata)

#ggcorrplot(corrdata, hc.order = TRUE,type = "lower", p.mat = p.mat)
```
## Scaling
At this point we fit data to a a range of between 0 and 1.

```{r  scaling, include=TRUE,echo = FALSE}
library(caret)
# vector of columns you DON'T want
#df1<- dataset1[,setdiff(names(dataset1), c("Total")), with = FALSE]# subset
#df1<-scale(df1)
#summary(df1)
```



```{r pca_Cluster, include=TRUE,echo = FALSE}
#PRINCIPAL COMPONENT ANALYSIS
#library(ggbiplot)
#residuals<- prcomp(df1,scale=T,center=T)

```



# T-DISTRIBUTED STOCHASTIC NEIGHBOR EMBEDDING(t-SNE)


```{r t-SNE, include=TRUE,echo = FALSE}
# Loading our tnse library
# 
library(Rtsne)
# Curating the database for analysis 
# 
Labels<-dataset1$Total
dataset1$Total<-as.factor(dataset1$Total)

# For plotting
#
colors = rainbow(length(unique(dataset1$Total)))
names(colors) = unique(dataset1$Total)

# Executing the algorithm on curated data
# 
tsne <- Rtsne(dataset1[,-1], dims = 2, perplexity=14, verbose=TRUE, max_iter = 200)

# Getting the duration of execution
# 
exeTimeTsne <- system.time(Rtsne(dataset1[,-1], dims = 2, perplexity=14, verbose=TRUE, max_iter = 200))

# Plotting our graph and closely examining the graph
# 
plot(tsne$Y, t='n', main="tsne")
text(tsne$Y, labels=dataset1$Total, col=colors[dataset1$Total])
```


# Part 2: Feature Selection

Importance of features can be estimated from data by building a Learning Vector Quantization (LVQ) model. The varImp is then used to estimate the variable importance, which is printed and plotted.

```{r RF, include=TRUE,echo = FALSE}
# load the library
library(randomForest)
library(caret)
# load the dataset
dataset2<-read.csv("http://bit.ly/CarreFourDataset",header =T)
data<-dataset2[,-c(1,2,3,4,5,8,9,11)]
#head(data)
# train the model
model <- randomForest(Total~., data=data)
# estimate variable importance
importance <- importance(model)
# summarize importance
importance
# plot importance
plot(importance)
```